Visualizing Contingency Tables

### 2006-12-12 trunk/vcd/paper/Shadings/JCGS.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@630 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<<load_data,echo=FALSE,results=hide>>=
library("vcd")

@@

<<inference,echo=FALSE,results=hide>>=
if(recompute) {
  set.seed(myseed)
  art_max <- coindep_test(art, n = nrep)
  set.seed(myseed)
  art_chisq <- coindep_test(art, n = nrep, indepfun = function(x) sum(x^2))

@@

<<auxiliary_functions,echo=FALSE,results=hide>>=
stat <- function(x, digits = 3) as.vector(round(x$statistic, digits = digits))
pval <- function(x, digits = 6) {
  oop <- options()
  on.exit(oop)
  options(scipen = 1000)
  as.vector(round(x$p.value, digits = digits))
}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\begin{center}
<<classic_plots,fig=TRUE,echo=FALSE,height=6,width=11,results=hide>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
pushViewport(viewport(layout.pos.col=1, layout.pos.row=1))
mosaic(art, newpage = FALSE, margins = c(2.5, 4, 2.5, 3))
popViewport()
pushViewport(viewport(layout.pos.col=2, layout.pos.row=1))
assoc(art, newpage = FALSE, margins = c(5, 2, 5, 4))
popViewport(2)
@
\caption{Classic mosaic and association plot for the arthritis data.}
\label{fig:classic}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<shading_HSV,echo=FALSE,results=hide>>=
png(file = "shading_HSV.png", height = 525, width = 600)
## generate colors
hue23 <- hue.slice(2/3, grid.n = 101, plot = FALSE, type = "HSV")
hue0 <- hue.slice(0, grid.n = 101, plot = FALSE, type = "HSV")
saturation <- as.numeric(colnames(hue23))
value <- as.numeric(rownames(hue23))

@@

<<shading_HCL,echo=FALSE,results=hide>>=
png(file = "shading_HCL.png", height = 525, width = 600)
## generate colors
hue260 <- hue.slice(260, grid.n = 101, plot = FALSE)
hue360 <- hue.slice(360, grid.n = 101, plot = FALSE)
mychroma <- as.numeric(colnames(hue260))
luminance <- as.numeric(rownames(hue260))

@@

\setkeys{Gin}{width=2\textwidth}
\begin{sidewaysfigure}[p]
\begin{center}
<<shadings,fig=TRUE,echo=FALSE,height=12,width=18,results=hide,include=FALSE>>=
##                                       width=16
mymar <- c(1.5, 0.5, 0.5, 2.5)
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 3)))
pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 1))
mosaic(art, newpage = FALSE, gp = shading_Friendly, margins = mymar,
  gp_args = list(eps = NULL, lty = 1))
popViewport()
pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 2))
mosaic(art, newpage = FALSE, gp = shading_hsv, margins = mymar,
  gp_args = list(interpolate = art_max$qdist(c(0.9, 0.99)), p.value = art_max$p.value))
popViewport()
pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 3))
mosaic(art, newpage = FALSE, gp = shading_hcl, margins = mymar,
  gp_args = list(interpolate = art_max$qdist(c(0.9, 0.99)), p.value = art_max$p.value))
popViewport()

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<conditional_mosaic,fig=TRUE,echo=FALSE,height=5,width=9>>=
cotabplot(~ Smoking + Disease | Gender, data = alz, panel = alz_cotab)
@
\caption{Conditional mosaic plot with double maximum shading for conditional
independence of smoking and disease given gender.}
\label{fig:alz}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\begin{center}
<<conditional_assoc,fig=TRUE,echo=FALSE,height=11,width=11>>=
cotabplot(~ Memory + Attitude | Age + Education, data = pun, panel = pun_cotab)
@
\caption{Conditional association plot with maximum sum-of-squares shading for conditional
independence of memory and attitude given age and education.}
\label{fig:pun}
\end{center}
\end{figure}

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\begin{center}
<<conditional_mosaic2,fig=TRUE,echo=FALSE,height=11,width=11>>=
cotabplot(~ Memory + Attitude | Age + Education, data = pun, panel = pun_cotab2)
@
\caption{Conditional mosaic plot with maximum sum-of-squares shading for conditional
independence of memory and attitude given age and education.}
\label{fig:pun2}
\end{center}
\end{figure}

>>>>

<<load_data>>=
library("vcd")

@@

<<inference>>=
if(recompute) {
  set.seed(myseed)
  art_max <- coindep_test(art, n = nrep)
  set.seed(myseed)
  art_chisq <- coindep_test(art, n = nrep, indepfun = function(x) sum(x^2))

@@

<<auxiliary_functions>>=
stat <- function(x, digits = 3) as.vector(round(x$statistic, digits = digits))
pval <- function(x, digits = 6) {
  oop <- options()
  on.exit(oop)
  options(scipen = 1000)
  as.vector(round(x$p.value, digits = digits))
}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\begin{center}
<<classic_plots,fig=TRUE,height=6,width=11>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
pushViewport(viewport(layout.pos.col=1, layout.pos.row=1))
mosaic(art, newpage = FALSE, margins = c(2.5, 4, 2.5, 3))
popViewport()
pushViewport(viewport(layout.pos.col=2, layout.pos.row=1))
assoc(art, newpage = FALSE, margins = c(5, 2, 5, 4))
popViewport(2)
@
\caption{Classic mosaic and association plot for the arthritis data.}
\label{fig:classic}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<shading_HSV>>=
png(file = "shading_HSV.png", height = 525, width = 600)
## generate colors
hue23 <- hue.slice(2/3, grid.n = 101, plot = FALSE, type = "HSV")
hue0 <- hue.slice(0, grid.n = 101, plot = FALSE, type = "HSV")
saturation <- as.numeric(colnames(hue23))
value <- as.numeric(rownames(hue23))

@@

<<shading_HCL>>=
png(file = "shading_HCL.png", height = 525, width = 600)
## generate colors
hue260 <- hue.slice(260, grid.n = 101, plot = FALSE)
hue360 <- hue.slice(360, grid.n = 101, plot = FALSE)
mychroma <- as.numeric(colnames(hue260))
luminance <- as.numeric(rownames(hue260))

@@

\setkeys{Gin}{width=2\textwidth}
\begin{sidewaysfigure}[p]
\begin{center}
<<shadings,fig=TRUE,height=12,width=18,include=FALSE>>=
##                                       width=16
mymar <- c(1.5, 0.5, 0.5, 2.5)
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 3)))
pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 1))
mosaic(art, newpage = FALSE, gp = shading_Friendly, margins = mymar,
  gp_args = list(eps = NULL, lty = 1))
popViewport()
pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 2))
mosaic(art, newpage = FALSE, gp = shading_hsv, margins = mymar,
  gp_args = list(interpolate = art_max$qdist(c(0.9, 0.99)), p.value = art_max$p.value))
popViewport()
pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 3))
mosaic(art, newpage = FALSE, gp = shading_hcl, margins = mymar,
  gp_args = list(interpolate = art_max$qdist(c(0.9, 0.99)), p.value = art_max$p.value))
popViewport()

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<conditional_mosaic,fig=TRUE,height=5,width=9>>=
cotabplot(~ Smoking + Disease | Gender, data = alz, panel = alz_cotab)
@
\caption{Conditional mosaic plot with double maximum shading for conditional
independence of smoking and disease given gender.}
\label{fig:alz}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\begin{center}
<<conditional_assoc,fig=TRUE,height=11,width=11>>=
cotabplot(~ Memory + Attitude | Age + Education, data = pun, panel = pun_cotab)
@
\caption{Conditional association plot with maximum sum-of-squares shading for conditional
independence of memory and attitude given age and education.}
\label{fig:pun}
\end{center}
\end{figure}

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\begin{center}
<<conditional_mosaic2,fig=TRUE,height=11,width=11>>=
cotabplot(~ Memory + Attitude | Age + Education, data = pun, panel = pun_cotab2)
@
\caption{Conditional mosaic plot with maximum sum-of-squares shading for conditional
independence of memory and attitude given age and education.}
\label{fig:pun2}
\end{center}
\end{figure}
